Flatiron data cleaning and analysis report

### 2021-03-12 Rmd/Report cleaning.Rmd
Finished Rmd explanation cleaning

```{r creatinine, rows.print=15, max.print=15}
creatinine %>% filter(str_detect(labcomponent, "blood|serum")) %>% 
arrange(testresultcleaned)
```
All `testresultcleaned` and `testresult` variable are cleaned as :  
-creat < 0.8 is set at 0.8  
-creat > 3 are eliminated  
To create the final `creatinine` variable ,`testresultcleaned` and `testresult` are both used but `testresultcleaned` is used over `testresult` when both are available.

@@

vitals <- vitals %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y"))
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)")
```
<br>

@@

`testresultcleaned` is all within expected range but only few data is available (`r NROW(body_surface_area %>% filter(!is.na(testresultcleaned)) %>% distinct(patientid, .keep.all=TRUE))` patients). Even less with an associated date!

```{r bsa}
body_surface_area
```

By using `testresult` and before cleaning only `r NROW(body_surface_area %>% filter(!is.na(body_surface_area$testresult)) %>% distinct(patientid, .keep.all=TRUE))` patients have data for their BSA. So I will also use the height and weight of patients to calculate their BSA by using the Du Bois formula.

To create the final `BSA` variable ,`testresultcleaned` is used over `testresult` when both are available.  
`testresult` was further cleaned as data < 0.418 and > 4.14 are eliminated. These values were calculated from the biggest and the smallest individuals recorded.  
I also remove the outliers within patient (2 examples here).

@@

Here a sample of the cleaned data

```{r bsa1, rows.print=15, max.print=15}
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)") %>% # only 399 patients data
  # Limit max = 4.14, min = 0.418 calculated for biggest to smallest person
  mutate(testresult = ifelse(testresult < 0.418 | testresult > 4.14, NA_real_, testresult)) %>% 
  group_by(patientid) %>%
  # I used the median to remove outlier within a patient
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(BSA = case_when(
    testresult > (median_testresult + 0.4) |
      testresult < (median_testresult - 0.4)      ~ NA_real_,
    TRUE                                          ~ testresult
  )) %>% 
  mutate(BSA = coalesce(testresultcleaned, BSA)) %>% 
  filter(!is.na(BSA)) %>% 
  mutate(BSA_units = "m2") %>% 
  select(c("patientid", bsa_date = "testdate", "BSA", "BSA_units"))

body_surface_area %>% distinct(patientid, .keep_all = TRUE)
```
<br>

@@

Height data complicated. Again you have `testresultcleaned` and `testresult` with `testresultcleaned` which will prioritized.  
`testresultcleaned` has units in cm but some data are wrong. `testresult` has results in cm, ft or in. It is pretty easy to clean data between these 3 units because the measurement cannot overlap.
Both are cleaned as < 120 or > 231 are eliminated. These value are for the shortest and tallest women recorded.
The units can be filled by patients when absent.

```{r height}
vitals %>% 
  filter(test == "body height" & patientid == "FCB6DAAFF6527")
# FC109A6F719E7
# FC106805F9AB9
# F91590CB83EC8 weird
```

@@

```{r height1}
vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  filter(patientid == "F01413C06D921") %>% 
  select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified,
           testunitscleaned, testunits, testresult))
```

PB: Some patient don't have dates but have data. Should I take a mean of all the height measured or the first one - If I have to choose I would take the mean (patients shouldn't grow or shorten much)?
If all dates are NA for a patients, I cannot combined `height` with `episodedate` of the drug data.

Here a samples of the data.
```{r height2, rows.print=15, max.print=15}
height <- vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  # select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified, 
  #          testunitscleaned, testunits, testresult)) %>%

@@

height %>% distinct(patientid, .keep_all = TRUE)
```
`r emo::ji("check")`

@@

The more complex cleaning (before the drugs...`r emo::ji("sad")`). Contrary to height with units which don't overlap. Weight has data in `kg` in `testresultcleaned` and `kg`, `lb`, `oz` in `testresult`which are overlapping (a person can be 150lbs or 150kg).  
Cleanup of `testresultcleaned`...  
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(patientid == "F0027D3926C88")
```
... and eliminate the outliers (36.4 kg can be normal for FF5268589DBE2, 39.5 F54F47F4419E0) 37.7 not F4A3B1B659865, F6CA354D599AA.
I have an idea on how to improve that better `r emo::ji("cross_mark")``r emo::ji("light_bulb")``r emo::ji("bookmark")`
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(str_detect(patientid, "F6CA354D599AA|F4A3B1B659865|FF5268589DBE2|F93CEAE4BF764|F1493F8F5A924|F18E68D19D10F|F54F47F4419E0"))
```

Cleanup of `testresult` for each units separately because can have 2 units per patients which can be correct or not.
Plus try to rescue the units when some are `NA` but not all and eliminate the outliers.
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% 
  # mutate(difference = (median_testresultcleaned)-testresultcleaned) # Can improve by > 30 and > 10 from previous measure so need to create a gap var ---------
  # 2.Clean testresult to be used to find max and min
  mutate(testresult = case_when(
    testunits == "lb" &
      testresult > 12 &
      testresult < 643            ~ testresult,
    testunits == "kg" &
      testresult >= 5.5 &
      testresult < 292            ~ testresult,
    testunits == "oz" &
      testresult > 194 &
      testresult < 10286          ~ testresult,
    is.na(testunits) & 
    testresult > 5.5 & 
      testresult < 10286          ~ testresult, # combine oz, kg, lbs for when units is NA
    TRUE                          ~ NA_real_
  )) %>% filter(str_detect(patientid, "F0027D3926C88"))

@@

Sometimes `testunits` is wrong or change so try to rescue the unit within the same patient.
Rescue units for when value is twice a `kg` value, it is `lb` value. 
```{r}
weight2 <- weight1 %>% 
  group_by(patientid) %>% 
  # 5. Rescue testunits when testresult_verified is twice or /2 means that lb or kg 
  # if is 2 times the weight in kg add lb F14EBD0D0FA81
  mutate(testunits_rescue1 = case_when(
    is.na(testunits) &
      testresult_verified < (max_kg * 2.205) &
      testresult_verified > (min_kg * 2.205)         ~ "lb",
    is.na(testunits) &
      testresult_verified < (max_lbs / 2.205) &
      testresult_verified > (min_lbs / 2.205)        ~ "kg"
  )) %>% 
  ungroup() %>% 
  mutate(testunits = coalesce(testunits, testunits_rescue, testunits_rescue1)) %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testresult_verified", "testunitscleaned", "testresultcleaned", "weight1"))
weight1 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
weight2 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
```

@@

F0027D3926C88
```{r}
weight <- weight2 %>% 
  # 6.Fill up weight with new testresult_verified cleaned
  mutate(weight2 = case_when(
    is.na(testresultcleaned) &
      testunits == "kg"         ~ testresult_verified, # F0027D3926C88, FE8DBCAC6DAFB, F8A3DD629CDA6, F020C6A8B9E50, F16847515CAF4
    is.na(testresultcleaned) & # F044601199C5D
      testunits == "lb"         ~ (testresult_verified / 2.205), # FCB918617347C, F5E3D88009911, F14EBD0D0FA81, F16847515CAF4, FABC474931881
  )) %>% 
  mutate(weight = coalesce(weight1, weight2)) %>% 
  filter(!is.na(weight)) %>% 
  mutate(weight_units = "kg") %>% 
  select(c("patientid", weight_date = "testdate", "weight", "weight_units"))
weight
```

@@

I selected only AUC collected for the taxel and platin drugs but can have really high and low number.
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(relativeorderedamount)
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(desc(relativeorderedamount))
```
I clean the data up by keeping the `relativeorderedamount`:  
- >= 50 & <= 175 for taxel  and with a unit in `mg/m2`  
- >= 2  & <= 6   for platin and with a unit in `AUC`
But few example for taxel
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  # select(c("patientid", "expectedstartdate", orderedamount, orderedunits, "relativeorderedamount",
  #          "relativeorderedunits", "drugname", iscanceled)) %>%
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    # relativeorderedunits == "mg/kg" |
    #   relativeorderedunits == "mg" |
    #   relativeorderedunits == "m"         ~ NA_real_,
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F21904E0D872F|F11B625154B62"))
```

But few example for platin
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  # select(c("patientid", "expectedstartdate", orderedamount, orderedunits, "relativeorderedamount",
  #          "relativeorderedunits", "drugname", iscanceled)) %>%
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    # relativeorderedunits == "mg/kg" |
    #   relativeorderedunits == "mg" |
    #   relativeorderedunits == "m"         ~ NA_real_,
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F4E8C8717878B|FE3FF3A5AC465|F853DC24CD5E8"))
```

>>>>

```{r creatinine, rows.print=15, max.print=15}
creatinine %>% filter(str_detect(labcomponent, "blood|serum")) %>% select(-c(loinc, test, labcomponent)) %>% 
arrange(testresultcleaned)
```
All `testresultcleaned` and `testresult` variable are cleaned as :  
- creat < 0.8 was set at 0.8  
- creat > 3 are eliminated  
To create the final `creatinine` variable ,`testresultcleaned` and `testresult` are both used but `testresultcleaned` was used over `testresult` when both are available.

@@

vitals <- vitals %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y"))
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)")
```
`r emo::ji("check")`
<br>

@@

`testresultcleaned` is all within expected range but only few data is available (`r NROW(body_surface_area %>% filter(!is.na(testresultcleaned)) %>% distinct(patientid, .keep.all=TRUE))` patients). Probably even less with an associated date.

```{r bsa}
body_surface_area %>% select(-c(loinc, test, labcomponent))
```

By using `testresult` and before cleaning only `r NROW(body_surface_area %>% filter(!is.na(body_surface_area$testresult)) %>% distinct(patientid, .keep.all=TRUE))` patients have data for their BSA. This was improved by using the Du Bois formula to calculate more BSA value as you will see later.

To create the final `BSA` variable ,`testresultcleaned` was used over `testresult` when both are available.  
`testresult` was further cleaned as data < 0.418 and > 4.14 are eliminated. These values were calculated from the biggest and the smallest individuals recorded.  
I also remove the outliers within patient (2 examples here where we have outliers).

@@

Here the cleaned data

```{r bsa1}
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)") %>% # only 399 patients data
  # Limit max = 4.14, min = 0.418 calculated for biggest to smallest person
  mutate(testresult = ifelse(testresult < 0.418 | testresult > 4.14, NA_real_, testresult)) %>% 
  group_by(patientid) %>%
  # I used the median to remove outlier within a patient
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(BSA = case_when(
    testresult > (median_testresult + 0.4) |
      testresult < (median_testresult - 0.4)      ~ NA_real_,
    TRUE                                          ~ testresult
  )) %>% 
  mutate(BSA = coalesce(testresultcleaned, BSA)) %>% 
  filter(!is.na(BSA)) %>% 
  mutate(BSA_units = "m2") %>% 
  select(c("patientid", bsa_date = "testdate", "BSA", "BSA_units"))

body_surface_area %>% distinct(patientid, .keep_all = TRUE) %>% arrange(patientid, bsa_date)
```
`r emo::ji("check")`
<br>

@@

Height data is pretty complicated. Again I used `testresultcleaned` and `testresult` with `testresultcleaned` which was prioritized.  
`testresultcleaned` has units in cm but some data are wrong. `testresult` has results in cm, ft or in. It was pretty easy to clean data between these 3 units because the measurement cannot overlap.
Both are cleaned as < 120 or > 231 are eliminated. These value are for the shortest and tallest women recorded.
The units can be filled by patients when absent.

```{r height}
vitals %>% 
  filter(test == "body height" & patientid == "FCB6DAAFF6527") %>% select(-c(loinc, test, labcomponent, resultdate))
# FC109A6F719E7
# FC106805F9AB9
# F91590CB83EC8 weird
```

@@

```{r height1}
vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  filter(patientid == "F01413C06D921") %>% 
  select(c("patientid", testdate, testresultcleaned, testresultcleaned_verified,median_testresultcleaned, 
           testunitscleaned, testunits, testresult))
```
`r  emo::ji("question_mark")`
PB: Some patients have height but don't have dates associated. If all dates are `NA` for a patient, I cannot combined a specific `height` with a closest date to `episodedate` of the drug data.
Should I take a mean of all the height values measured or the first one - If I have to choose I would take the mean (patients shouldn't grow up or shorten much)?

Here is the data.
```{r height2}
height <- vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  # select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified, 
  #          testunitscleaned, testunits, testresult)) %>%

@@

height %>% distinct(patientid, .keep_all = TRUE)
```
Beside my question it is `r emo::ji("check")`

@@

The more complex cleaning (before the drugs...`r  emo::ji("loudly_crying_face")`). Contrary to height with units which don't overlap, weight has data in `kg` in `testresultcleaned` and `kg`, `lb`, `oz` in `testresult` which are overlapping (a person can be 150lbs or 150kg).  
Cleanup of `testresultcleaned`...  
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(patientid == "F0027D3926C88")
```
... and eliminate the outliers (a patient gaining 36.4 kg can be normal for FF5268589DBE2, 39.5kg F54F47F4419E0) but 37.7 doesn't look normal F4A3B1B659865, F6CA354D599AA. My present cleaning save some data but not everything...  
I just had an idea on how to improve that better while doing the report.  
`r emo::ji("cross_mark")``r emo::ji("light_bulb")``r emo::ji("bookmark")`
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(str_detect(patientid, "F6CA354D599AA|F4A3B1B659865|FF5268589DBE2|F93CEAE4BF764|F1493F8F5A924|F18E68D19D10F|F54F47F4419E0"))
```

I had to clean `testresult` for each units separately within a patients because I saw patients with 2 units in their records.
I also rescued the units when some are `NA` but not all and eliminate the outliers.

```{r}
# vitals %>%
#   filter(test == "body weight") %>% 
#   select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
#   # 1.Clean testresultcleaned
#   group_by(patientid) %>%
#   mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
#   ungroup() %>%
#   mutate(weight1 = case_when(
#     testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
#       testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
#       (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
#          testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
#     TRUE ~ NA_real_
#   )) %>% 
#   # mutate(difference = (median_testresultcleaned)-testresultcleaned) # Can improve by > 30 and > 10 from previous measure so need to create a gap var ---------
#   # 2.Clean testresult to be used to find max and min
#   mutate(testresult = case_when(
#     testunits == "lb" &
#       testresult > 12 &
#       testresult < 643            ~ testresult,
#     testunits == "kg" &
#       testresult >= 5.5 &
#       testresult < 292            ~ testresult,
#     testunits == "oz" &
#       testresult > 194 &
#       testresult < 10286          ~ testresult,
#     is.na(testunits) & 
#     testresult > 5.5 & 
#       testresult < 10286          ~ testresult, # combine oz, kg, lbs for when units is NA
#     TRUE                          ~ NA_real_
#   )) %>% filter(str_detect(patientid, "F0027D3926C88")) %>% 
#   select(c("patientid", "testdate", "testunits", "testresult",)

@@

Sometimes `testunits` was wrong or changed so try to rescue the unit within the same patient. For example these 2 patients have data in `kg` and earlier data with no units but doubled values so I assumed these data where on `lb.`

```{r}
weight2 <- weight1 %>% 
  group_by(patientid) %>% 
  # 5. Rescue testunits when testresult_verified is twice or /2 means that lb or kg 
  # if is 2 times the weight in kg add lb F14EBD0D0FA81
  mutate(testunits_rescue1 = case_when(
    is.na(testunits) &
      testresult_verified < (max_kg * 2.205) &
      testresult_verified > (min_kg * 2.205)         ~ "lb",
    is.na(testunits) &
      testresult_verified < (max_lbs / 2.205) &
      testresult_verified > (min_lbs / 2.205)        ~ "kg"
  )) %>% 
  ungroup() %>% 
  mutate(testunits = coalesce(testunits, testunits_rescue, testunits_rescue1)) %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testresult_verified", "testunitscleaned", "testresultcleaned", "weight1"))
weight1 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
```

The units are rescued here.
```{r}
weight2 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
```

@@

<!-- F0027D3926C88 -->
```{r}
weight <- weight2 %>% 
  # 6.Fill up weight with new testresult_verified cleaned
  mutate(weight2 = case_when(
    is.na(testresultcleaned) &
      testunits == "kg"         ~ testresult_verified, # F0027D3926C88, FE8DBCAC6DAFB, F8A3DD629CDA6, F020C6A8B9E50, F16847515CAF4
    is.na(testresultcleaned) & # F044601199C5D
      testunits == "lb"         ~ (testresult_verified / 2.205), # FCB918617347C, F5E3D88009911, F14EBD0D0FA81, F16847515CAF4, FABC474931881
  )) %>% 
  mutate(weight = coalesce(weight1, weight2)) %>% 
  filter(!is.na(weight)) %>% 
  mutate(weight_units = "kg") %>% 
  select(c("patientid", weight_date = "testdate", "weight", "weight_units"))
weight
```

I can probably improve my cleaning the same way as height but beside that is it `r emo::ji("check")`
<br>

@@

I selected only AUC collected for the "taxel" and "platin" drugs but can have really high and low values.
```{r, rows.print=15, max.print=15}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(relativeorderedamount) %>% select(patientid, relativeorderedamount, orderedamount, orderedunits, everything())
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(desc(relativeorderedamount)) %>% select(patientid, relativeorderedamount, orderedamount, orderedunits, everything())
```
I clean the data up by keeping the `relativeorderedamount`:  
- >= 50 & <= 175 for taxel  and with a unit in `mg/m2`  
- >= 2  & <= 6   for platin and with a unit in `AUC`
`r  emo::ji("question_mark")` But few example for taxel are remove for a `relativeorderedamount` at 36. Should we open it up a little more?
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F21904E0D872F|F11B625154B62")) %>% select(patientid, drugname, relativeorderedamount,relativeorderedunits, target_auc, orderedamount, orderedunits, expectedstartdate)
```

`r  emo::ji("question_mark")` Same for platin?
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F4E8C8717878B|FE3FF3A5AC465|F853DC24CD5E8")) %>% select(patientid, drugname, relativeorderedamount,relativeorderedunits, target_auc, orderedamount, orderedunits, expectedstartdate)
```
